---
title: "Correlations and multicollinearity"
output: html_notebook
---

This script checks for mutlicollinearity and correlations across all features. This needs to be removed in order to perform a multinomial regression.

First load the features:

```{r}
features = read.csv("combined_features.csv")
```

Now create a function to create the heatmaps
```{r}
create_correlation_heatmap = function(df, labels, NA_cols = c(), title = NULL, type = "correlation"){
  NA_cols = c(NA_cols, "filename") # also remove filename for the correlation matrix
  df = df[, !names(df) %in% NA_cols]
  print(paste(nrow(df) - nrow(na.omit(df)), "rows were removed"))
  if (type == "correlation"){
    corr_df = cor(na.omit(df))
  } else if (type == "partial_correlation"){
    library(corpcor)
    corr_df = as.data.frame(cor2pcor(cov(na.omit(df)))) # Compute Partial Correlation from Correlation Matrix
    row.names(corr_df) = names(df)
    names(corr_df) = names(df)
  } else {
    stop("Type not supported")
  }
  library(rstatix)
  corr_df = corr_df %>% cor_gather()
  corr_df$cor = abs(corr_df$cor)
  
  if (length(labels) != ncol(df)){
    stop("Same number of labels as names in DF!")
  }
  
  # Make them to factors, so they don't change position
  corr_df$var1 = factor(corr_df$var1, levels = names(df))
  corr_df$var2 = factor(corr_df$var2, levels = names(df))
  
  colors = c("#00bd9d", "#12ce76", "#2f96d9", "#9d54b2", "#f2c62f", "#e97f2c", "#eb4c3e")
  
  if (length(unique(labels)) > length(colors)){
    stop("To many labels")
  } else{
    colors = colors[1:length(unique(labels))]
  }
  
  
  names = unique(corr_df$var1)
  names = factor(names, levels = names)
  
  fill = factor(labels, levels = unique(labels))
  #fill = labels
  
  heatmap = ggplot(corr_df, aes(y=var1, x=var2, fill=cor)) + 
    labs(x = "features", y = "features", fill = "Absolute correlation") +
    geom_point(aes(color=factor(rep(fill, length(fill)), levels = unique(fill))), alpha=0) +
    geom_tile(colour="white") + 
    theme_minimal() + 
    theme(
      axis.title.y = element_blank(),
      axis.title.x = element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks.y=element_blank(),
      panel.border=element_blank(),
      panel.grid=element_blank(),
      plot.margin=unit(c(0, 0,0,0),"cm")
    ) + 
    scale_color_manual("Measures", values=colors, drop=FALSE) +
    scale_fill_gradient(low = "white", high = "black") +
    guides(color = guide_legend(override.aes = list(alpha=1)))

  if (is.null(title)){
    padding_top = 10
    padding_left = 10
  } else{
    heatmap = heatmap +  ggtitle(title)
    padding_top = 20
    padding_left = 15
  }
  
  y_annotate = ggplot(data.frame(y = names, x = 0, fill = fill), aes(y=y, x =x, fill=fill)) + 
    geom_tile(colour="white") +
    cowplot::theme_nothing() +
    theme(
      axis.title = element_blank(),
      axis.text = element_blank(),
      legend.position = "none",
      panel.border=element_blank(),
      panel.grid=element_blank(),
      plot.margin=unit(c(0,.5,0,0),"cm")
    ) + 
    scale_fill_manual(values=colors)
    
  
  x_annotate = ggplot(data.frame(x = names, y = 0, fill = fill), aes(x=x, y =y, fill=fill)) + 
    geom_tile(colour="white") +
    cowplot::theme_nothing() +
    theme(
      axis.title = element_blank(),
      axis.text = element_blank(),
      legend.position = "none",
      panel.border=element_blank(),
      panel.grid=element_blank(),
      plot.margin=unit(c(0, 4.2, 0, 0.1),"cm")
    ) +
    coord_cartesian(ylim = c(0, 0)) +
    scale_fill_manual(values=colors)
  
  print(cowplot::plot_grid(y_annotate, heatmap, NULL, x_annotate, align = 'h', rel_widths = c(1, padding_top), rel_heights = c(10, 1)))
}
```

Some variables contain many NAs. How many columns remain, if we remove them?
```{r}
library(tidyverse)
freq_df = data.frame(feature_name = names(features), num_NAs = as.numeric(map(features, ~sum(is.na(.)))))
freq_df[order(freq_df$num_NAs, decreasing = TRUE), ]

```

Create a heatmap of the correlations between the variables and the partial correlations
```{r}
new_ft_idxs = 90:151
labels = c(rep("shape", 40), rep("slope", 15), rep("distribution", 3))
#create_correlation_heatmap(features[,new_ft_idxs], labels, NA_cols, paste("Correlations without", paste(NA_cols, collapse = ", ")))
create_correlation_heatmap(features[,new_ft_idxs], labels, NA_cols, "Correlations between new features")
```

```{r}
create_correlation_heatmap(features[,new_ft_idxs], labels, NA_cols, "Partial correlations between new features", type="partial_correlation")

```


Which variables correlate a lot?
```{r}
corr_df = cor(na.omit(features[, !names(features) %in% c(NA_cols, "filename")]))
bool_mat = as.data.frame(abs(corr_df) > 0.5)
features_high_corr = data.frame(features = names(bool_mat)[new_ft_idxs], correlation = ((as.numeric(map(bool_mat, ~sum(.)))[new_ft_idxs])))

# Count the correlations:

ggplot(features_high_corr) + 
  geom_bar(aes(x=features, y=correlation, group=features), stat="identity", fill="#1cb89b") + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  ggtitle(paste0("Counts absolute correlations > .5 (counts with maximum is total features - 1: ", ncol(bool_mat) - 1,")"))
```




Correlations across baseline + new features
```{r}
labels = c(rep("eGeMAPS frequency", 30), rep("eGeMAPS energy", 15), rep("eGeMAPS spectrum", 37), rep("eGeMAPS duration", 6), rep("shape", 40), rep("slope", 15), rep("distribution", 3))

label_lookup = data.frame(labels = labels, features = names(features)[!names(features) %in% c("filename", NA_cols)])

create_correlation_heatmap(features, labels, NA_cols = NA_cols, title = "Correlations between variables in corpus")
```

```{r}
create_correlation_heatmap(features, labels, NA_cols = NA_cols, type = "partial_correlation", title = "Partial correlation")
```

```{r}
features_to_inspect = features[, !names(features) %in% c(NA_cols, "emotion", "filename")]
emotion = as.numeric(meta_df$emotion)
class_df = as.data.frame(sapply(features_to_inspect, class))
class_df$var = row.names(class_df)
names(class_df)[1] = "class"

for (col in dplyr::filter(class_df, class=="integer")$var){
  features_to_inspect[[col]] = as.numeric(features_to_inspect[[col]])
}
library(mctest)
X = as.matrix(features_to_inspect)
omcdiag(X, emotion)
```


```{r}
results = imcdiag(X,emotion)
vif_table = as.data.frame(results$idiags)
vif_table$feature = row.names(vif_table)
vif_table = vif_table[order(vif_table$VIF, decreasing = TRUE), ]
```


```{r}
imcdiag(X[,  names(features_to_inspect) %in% c(vif_table[140:146, "feature"], "F0semitoneFrom27.5Hz_sma3nz_amean", "F0semitoneFrom27.5Hz_sma3nz_stddevNorm", "loudness_sma3_amean", "VoicedSegmentsPerSec", "hammarbergIndexV_sma3nz_amean", "alphaRatioV_sma3nz_amean", "slopeV500.1500_sma3nz_amean")], emotion)
imcdiag(X[,  names(features_to_inspect) %in% vif_table[126:146, "feature"]], emotion)
```

Mutlicolinearity in Banse & Scherer features:

```{r}
imcdiag(X[,  names(features_to_inspect) %in% c("F0semitoneFrom27.5Hz_sma3nz_amean", "F0semitoneFrom27.5Hz_sma3nz_stddevNorm", "loudness_sma3_amean", "VoicedSegmentsPerSec", "hammarbergIndexV_sma3nz_amean", "alphaRatioV_sma3nz_amean", "slopeV500.1500_sma3nz_amean")], emotion)
imcdiag(X[,  names(features_to_inspect) %in% c("F0semitoneFrom27.5Hz_sma3nz_amean", "F0semitoneFrom27.5Hz_sma3nz_stddevNorm", "loudness_sma3_amean", "VoicedSegmentsPerSec", "hammarbergIndexV_sma3nz_amean", "slopeV500.1500_sma3nz_amean")], emotion)
```



```{r}
ft_glm = features_to_inspect
ft_glm$emotion = meta_df$emotion
library(randomForest)
fit_rf = randomForest(emotion~., data=na.omit(ft_glm))
important_params = as.data.frame(importance(fit_rf))
important_params$features = row.names(important_params)
row.names(important_params) = NULL
important_params = important_params[order(important_params$MeanDecreaseGini, decreasing = TRUE), ]
head(important_params, n = 50)
```

```{r}
features_by_importance = important_params$features

# initially include all
vif_table = as.data.frame(imcdiag(X[,  features_by_importance], emotion)$idiags)
vif_table$feature = row.names(vif_table)
vif_table = vif_table[order(vif_table$VIF, decreasing = TRUE), ]

continue = TRUE

while (continue){
  print(paste("remove", vif_table$feature[1]))
  features_by_importance = features_by_importance[features_by_importance != vif_table$feature[1]]
  vif_table = as.data.frame(imcdiag(X[,  names(features_to_inspect) %in% features_by_importance], emotion)$idiags)
  vif_table$feature = row.names(vif_table)
  vif_table = vif_table[order(vif_table$VIF, decreasing = TRUE), ]
  
  if (!length(which(vif_table$Klein == 1))  > 3){
    print("Model converged")
    continue = FALSE
  }
}

#imcdiag(X[,  names(features_to_inspect) %in% important_params$features], emotion)


```

```{r}

```

